### 2013-01-19 edupar-13/410.tex
Fixed SIMFLEX citation.
410.tex: Compress wording in Pebbles syscall table.

As mentioned above, the main execution and debugging
platform used in the class is
Simics~\cite{simics},
originally developed by the Swedish Institute of
Computer Science and
currently a product of Wind River.
Unlike some emulators,
which focus on fast execution of correct code,
Simics
provides very faithful bit-level support
not only for code that behaves correctly but
also for code that is ``incorrect within reason.''
Occasionally student code manages to invoke an unsupported
function of some hardware device or even to cause the
simulator to crash, but this is quite rare, and in almost
every case when the ``bad event'' is characterized it turns
out that what the software requested from the hardware was
a mistake.
Meanwhile, inside its large envelope, Simics supports
\textit{very} detailed models of various x86 processors.
Unlike hardware virtualization environments,
Simics contains substantial debugger support:
single-stepping,
printing of source-level symbolic
expressions,
stack tracing,
display of TLB entries,
and even summaries of x86 hardware-defined descriptor tables.
Unlike some symbolic debuggers,
Simics is scriptable in Python,
and it is possible to interpose user-written measurement
frameworks to evaluate proposed hardware
%% TODO Fix citations. But also, maybe this mention isn't relevant and
%% could be left out (if we're in space trouble?) -- bblum
features~\cite{UW-GEMS,SimFlex,FeS2}.
Performance is very reasonable (Simics can simulate certain
cases, such as a CPU halted awaiting a clock interrupt,
much faster than real time).
%% TODO: One huge advantage of simics over QEMU that we should mention is
%% that QEMU only issues timer interrupts at basic block boundaries. -- bblum

@@

\begin{figure}
	\center \footnotesize
	\begin{tabular}{|l|p{0.28\textwidth}|}
		\hline
		\bf System call name & \bf Summary \\
		\hline
		\multicolumn{2}{c}{\em Lifecycle management} \\
		\hline
		\x{fork} & Duplicates the invoking task, including all memory regions. \\
		\x{thread_fork} & Creates a new thread in the current task.\\
		\x{exec} & Replaces the program currently running in the invoking task with a new one specified. \\
		\x{set_status} & Records the exit status of the current task. \\
		\x{vanish} & Terminates execution of the calling thread. \\
		\x{wait} & Blocks execution until another task terminates, and collects its exit status.\\
		\x{task_vanish}* & Causes all threads of a task to \x{vanish}. \\
		\hline
		\multicolumn{2}{c}{\em Thread management} \\
		\hline
		\x{gettid} & Returns the ID of the invoking thread. \\
		\x{yield} & Defers execution to a specified thread. \\
		\x{deschedule} & Blocks execution of the invoking thread. \\
		\x{make_runnable} & Wakes up another \x{deschedule}d thread. \\
		\x{get_ticks} & Gets the number of timer ticks since bootup. \\
		\x{sleep} & Blocks a thread for a given number of ticks. \\
		\x{swexn} & Registers a given user-space function as a software exception handler.\\
		\hline
		\multicolumn{2}{c}{\em Memory management} \\
		\hline
		\x{new_pages} & Allocates a specified region of memory. \\
		\x{remove_pages} & Deallocates same. \\
		\hline
		\multicolumn{2}{c}{\em Console I/O} \\
		\hline
		\x{getchar}* & Reads one character from keyboard input. \\
		\x{readline} & Reads the next line from keyboard input. \\
		\x{print} & Prints a given memory buffer to the console. \\
		\x{set_term_color} & Sets the color for future console output. \\
		\x{set_cursor_pos} & Sets the console cursor location. \\
		\x{get_cursor_pos} & Retrieves the console cursor location. \\
		\hline
		\multicolumn{2}{c}{\em Miscellaneous} \\
		\hline
		\x{ls} & Loads a given buffer with the names of files stored in the RAM disk ``file system''. \\
		\x{halt} & Ceases execution of the operating system. \\
		\x{misbehave}* & Selects among several behaviors for thread interleaving in certain scenarios. \\
		\hline
	\end{tabular}
	\caption{The Pebbles specifcation defines 25 system calls. Students are not required to implement ones marked with an asterisk (*), though the reference kernel provides them. }
	\label{fig:syscalls}
\end{figure}

>>>>

As mentioned above, the main execution and debugging
platform used in the class is
Simics~\cite{simics},
originally developed by the Swedish Institute of
Computer Science and
currently a product of Wind River.
Unlike some emulators,
which focus on fast execution of correct code,
Simics
provides very faithful bit-level support
not only for code that behaves correctly but
also for code that is ``incorrect within reason.''
Occasionally student code manages to invoke an unsupported
function of some hardware device or even to cause the
simulator to crash, but this is quite rare, and in almost
every case when the ``bad event'' is characterized it turns
out that what the software requested from the hardware was
a mistake.
Meanwhile, inside its large envelope, Simics supports
\textit{very} detailed models of various x86 processors.
Unlike hardware virtualization environments,
Simics contains substantial debugger support:
single-stepping,
printing of source-level symbolic
expressions,
stack tracing,
display of TLB entries,
and even summaries of x86 hardware-defined descriptor tables.
Unlike some symbolic debuggers,
Simics is scriptable in Python,
and it is possible to interpose user-written measurement
frameworks to evaluate proposed hardware
%% TODO Fix citations. But also, maybe this mention isn't relevant and
%% could be left out (if we're in space trouble?) -- bblum
features~\shortversion{\cite{SIMFLEX}}{\cite{SIMFLEX,UW-GEMS,FeS2}}.
Performance is very reasonable (Simics can simulate certain
cases, such as a CPU halted awaiting a clock interrupt,
much faster than real time).
%% TODO: One huge advantage of simics over QEMU that we should mention is
%% that QEMU only issues timer interrupts at basic block boundaries. -- bblum

@@

\begin{figure}
	\center \footnotesize
	\begin{tabular}{|l|p{0.28\textwidth}|}
		\hline
		\bf System call name & \bf Summary \\
		\hline
		\multicolumn{2}{c}{\em Lifecycle management} \\
		\hline
		\x{fork} & Duplicates the invoking task, including all memory regions. \\
		\x{thread_fork} & Creates a new thread in the current task.\\
		\x{exec} & Replaces the program currently running in the invoking task with a new one specified. \\
		\x{set_status} & Records the exit status of the current task. \\
		\x{vanish} & Terminates execution of the calling thread. \\
		\x{wait} & Blocks execution until another task terminates, and collects its exit status.\\
		\x{task_vanish}* & Causes all threads of a task to \x{vanish}. \\
		\hline
		\multicolumn{2}{c}{\em Thread management} \\
		\hline
		\x{gettid} & Returns the ID of the invoking thread. \\
		\x{yield} & Defers execution to a specified thread. \\
		\x{deschedule} & Blocks execution of the invoking thread. \\
		\x{make_runnable} & Wakes up another \x{deschedule}d thread. \\
		\x{get_ticks} & Gets the number of timer ticks since bootup. \\
		\x{sleep} & Blocks a thread for a given number of ticks. \\
		\x{swexn} & Registers a user-space software exception handler.\\
		\hline
		\multicolumn{2}{c}{\em Memory management} \\
		\hline
		\x{new_pages} & Allocates a specified region of memory. \\
		\x{remove_pages} & Deallocates a memory region. \\
		\hline
		\multicolumn{2}{c}{\em Console I/O} \\
		\hline
		\x{getchar}* & Reads one character from keyboard input. \\
		\x{readline} & Reads the next line from keyboard input. \\
		\x{print} & Prints a given memory buffer to the console. \\
		\x{set_term_color} & Sets the color for future console output. \\
		\x{set_cursor_pos} & Sets the console cursor location. \\
		\x{get_cursor_pos} & Retrieves the console cursor location. \\
		\hline
		\multicolumn{2}{c}{\em Miscellaneous} \\
		\hline
		\x{ls} & Loads a given buffer with the names of files stored in the RAM disk ``file system.'' \\
		\x{halt} & Ceases execution of the operating system. \\
		\x{misbehave}* & Selects among several thread-scheduling policies. \\
		\hline
	\end{tabular}
	\caption{The Pebbles specifcation defines 25 system calls. Students are not required to implement ones marked with an asterisk (*), though the reference kernel provides them. }
	\label{fig:syscalls}
\end{figure}
